---
title: "Stat442 project R scripts"
author: "Jianing Ren"
date: "05/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 0. Loading pkgs and dataset; don't run this chunk

```{r, cache=T}
library(airportr)
library(ggplot2)
library(dplyr)
library(rvest)
library(gam)
library(leaps)

air = read.csv("/Users/rjn/OneDrive/大学/spr_2021/stat442/442project/stat442-project/data/JAN2019 entire data wo divert.csv")
plane = read.csv("/Users/rjn/OneDrive/大学/spr_2021/stat442/442project/stat442-project/data/ReleasableAircraft.2019/MASTER.txt")  #FAA database containing aircraft model info 
weather = read.csv("/Users/rjn/OneDrive/大学/spr_2021/stat442/442project/stat442-project/weather.csv")

```

# 1. Data cleaning works 
## 1a. Filtering to desired rows

```{r}
list_of_airlines = c("AA","DL","UA")
list_of_airports = c("JFK","LAX","DEN","DFW","ORD","ATL")

air_filtered = air %>% filter(OP_UNIQUE_CARRIER %in% list_of_airlines
                      & ORIGIN %in% list_of_airports
                      & DEST %in% list_of_airports)
air_filtered$TAIL_NUM = as.character(air_filtered$TAIL_NUM)
```

## 1b. Extract and merge the aircraft model info from FAA database

```{r}
plane_sub = plane %>% select(N.NUMBER,MFR.MDL.CODE)

air_filtered$N.NUMBER = substring(air_filtered$TAIL_NUM,2,6)

air_combined = merge(air_filtered, plane_sub, by.y="N.NUMBER")  #removes flights that fails to match (either they don't have a tail number or their tail number isn't recorded in the FAA db). 
```

## 1c. Get manufacturer data

```{r}
air_combined$MFR=substring(air_combined$MFR.MDL.CODE,1,2)

#39 is Airbus; 30 is MD; 13 is Boeing
dict = data.frame(code=c(39,30,13), type=c("Airbus", "McDonnel-Douglas", "Boeing"))
air_combined$MFR.Actual = dict[match(air_combined$MFR, dict$code),]$type
```

## 1d. Route
```{r}
air_combined$route = paste0(as.character(air_combined$ORIGIN), "-",as.character(air_combined$DEST))
```


## 1f. Drop unused factor levels
```{r}
for (i in 1:ncol(air_combined))
{
    if (class(air_combined[[i]])=="factor")
        air_combined[[i]] = factor(air_combined[[i]])
}
```

## 1g. Merging Weather Data
```{r}
weather.table = data.frame(NAME = levels(weather$NAME), code=c("ATL","ORD","DFW","DEN","JFK","LAX"))
weather = merge(weather,weather.table, by.y="NAME")

origin.weather = weather[,c(19,4:8)]
names(origin.weather)[3:6] = paste0("origin_",names(origin.weather)[3:6])

dest.weather = weather[,c(19,4:8)]
names(dest.weather)[3:6] = paste0("dest_",names(dest.weather)[3:6])

air_with_weather = merge(air_combined, origin.weather, by.x=c("FL_DATE","ORIGIN"), by.y=c("DATE","code"))
air_with_weather = merge(air_with_weather, dest.weather, by.x=c("FL_DATE","DEST"), by.y=c("DATE","code"))

#write.csv(air_with_weather, "air_with_weather.csv")

```


```{r}
#LAX has missing snow data; convert them to zero

ori.index = which(air_with_weather$ORIGIN=="LAX")
air_with_weather[ori.index, ]$origin_SNOW=0

dest.index = which(air_with_weather$DEST=="LAX")
air_with_weather[dest.index,]$dest_SNOW=0
```


## 1h. Get airport coordinates (remove if unnecessary)

```{r}

locations = as.data.frame(sapply(list_of_airports, airport_location))
locations = apply(locations, 1, unlist)
locations = data.frame(airports = rownames(locations), locations)


air_with_weather = merge(air_with_weather, locations, by.x=("ORIGIN"), by.y=("airports"))

names(air_with_weather)[c(length(names(air_with_weather))-1, length(names(air_with_weather)))] = c("origin_lat", "origin_long")

air_with_weather = merge(air_with_weather, locations, by.x=("DEST"), by.y=("airports"))

names(air_with_weather)[c(length(names(air_with_weather))-1, length(names(air_with_weather)))] = c("dest_lat", "dest_long")
```


# 2.EDA
```{r}
air_for_plotting = air_with_weather[,-c(3,4,5,6,10,11,33:41)]
air_for_plotting2 = air_for_plotting[,c(17,31:42)]

plot(log(ARR_DELAY)~., data=air_for_plotting2)

#air_for_plotting = air_for_plotting %>% gather(-ARR_DELAY, key ="var", value="val")

boxplot(ARR_DELAY~OP_UNIQUE_CARRIER, data=air_combined)
boxplot(ARR_DELAY~MFR.Actual, data=air_combined)
boxplot(ARR_DELAY~route, data=air_combined)
plot(log(ARR_DELAY)~CRS_ARR_TIME, data=air_combined)

plot(log(ARR_DELAY)~origin_AWND, data=air_with_weather)

```

## Delay map

```{r}
states = map_data("state")

blank.map = ggplot() + 
    geom_polygon(data=states, 
                 aes(x=long, y=lat, group=region),
                 fill="#8eb6ed", color="white")


```

# 3. Modelling
## 3a. Preparing data
```{r}
# use testing from 2020.1 data? 

# removing variables unsuitable for modelling 

# how to deal with the reason of delay? 

data.for.modelling = air_with_weather %>% select(-N.NUMBER,-YEAR,-MONTH,-TAIL_NUM,-OP_CARRIER_FL_NUM,-DEP_TIME,-DEP_DELAY_GROUP,-ARR_TIME,-CANCELLATION_CODE,-CANCELLED,-DIVERTED,-ACTUAL_ELAPSED_TIME,-FLIGHTS, -DISTANCE_GROUP, -CARRIER_DELAY,-WEATHER_DELAY,-NAS_DELAY,-SECURITY_DELAY,-LATE_AIRCRAFT_DELAY,-FIRST_DEP_TIME,-TOTAL_ADD_GTIME,-LONGEST_ADD_GTIME, -X,-MFR, -route, -origin_lat, -origin_long, -dest_lat, -dest_long, -WHEELS_OFF, -WHEELS_ON)


#? ONLY USE PARAMETERS MEASURED BEFORE TAKING OFF? i.e., include taxi in/out or departure delay or not
data.for.modelling = data.for.modelling %>% select(-DEP_DELAY,-DEP_DELAY_NEW,-TAXI_IN, -TAXI_OUT)


#classifying departure and arrival time

time_breaks = c(0,900,1200,1400,1800,2100,2359)
time_labels = c("early morning","late morning","early afternoon","late afternoon","early night","late night")

data.for.modelling$CRS_DEP_TIME_GROUP = cut(data.for.modelling$CRS_DEP_TIME, breaks=time_breaks,labels=time_labels)
data.for.modelling$CRS_ARR_TIME_GROUP = cut(data.for.modelling$CRS_ARR_TIME, breaks=time_breaks,labels=time_labels)

data.for.modelling = data.for.modelling %>% select (-CRS_ARR_TIME,-CRS_DEP_TIME)

#converting day of week to actual name
data.for.modelling$DAY_OF_WEEK = as.factor(weekdays(as.Date(data.for.modelling$FL_DATE)))
data.for.modelling = data.for.modelling %>% select (-FL_DATE) #won't be necessary given day of month


```

## Further processing on missing/incomplete data
```{r}
#need to drop unique MFR.MDL.CODE, at least there shouldn't be new ones in testing. Therefore, filter out those plane types whose instances are fewer than 10

data.for.modelling = data.for.modelling %>% group_by(MFR.MDL.CODE) %>% filter(n()>10)
data.for.modelling$MFR.MDL.CODE = factor(data.for.modelling$MFR.MDL.CODE)

# remove other nas

data.for.modelling = data.for.modelling[complete.cases(data.for.modelling),]

#split training and testing
set.seed(1)
randvec = sample.int(nrow(data.for.modelling), 8000)
train2019 = data.for.modelling[randvec,]
test2019 = data.for.modelling[-randvec,]


table(test2019$MFR.MDL.CODE)
table(train2019$MFR.MDL.CODE)
```


## 3b. LM
```{r}
#AIR_TIME is probably perfectly correlated with flight distance and CRS_ELAPSED_TIME, therefore only keep one of them. If include more than one of them, R^2 will be close to 1

lm1 = lm(ARR_DELAY~.-CRS_ELAPSED_TIME-AIR_TIME-MFR.Actual, data=train2019)

#summary(lm1)

pred.lm = predict(lm1, test2019)
mean((pred.lm - test2019$ARR_DELAY)^2)
```

lm accuracy is plus/minus sqrt(1671) = 40.8 minutes, isn't quite good yet

## 3c. regsubsets [NOT WORKING]

```{r}
# not working; "really big"? 

lm.step = regsubsets(ARR_DELAY~.-CRS_ELAPSED_TIME-AIR_TIME-MFR.Actual, data=train2019, nvmax=2)
```

## 3d. lasso
```{r}
set.seed(1)

lambda.grid = 10^seq(-5,5, length.out=50)
train.matrix = model.matrix(ARR_DELAY~.-CRS_ELAPSED_TIME-AIR_TIME-MFR.Actual, data=train2019)

test.matrix = model.matrix(ARR_DELAY~.-CRS_ELAPSED_TIME-AIR_TIME-MFR.Actual, data=test2019)

air.lasso = cv.glmnet(train.matrix, train2019$ARR_DELAY, alpha=1, lambda=lambda.grid, nfolds=10)

plot(air.lasso)
coef.min = coef(air.lasso, s="lambda.min")

lasso.pred = predict(air.lasso, newx = test.matrix, s="lambda.min")
mean((lasso.pred - test2019$ARR_DELAY)^2)
```

## 3e. ridge (for correlated predictors)

## 3f. elastic net

# 99. Dumpster (testing or failed codes) 

### test in merging
```{, eval=F}
df1=data.frame(N.Number=c(1,3,5,6),rand=c(1000,1001,1002,1003))
df2=data.frame(N.Number=c(1,2,3,4,5), plane=c(717,727,737,747,757))
merge(df1,df2,by.y="N.Number",all.x=T)
```


### scraping

```{, eval=F}
url = paste0("https://flightaware.com/live/flight/", "N839AA")
website = read_html(url)
nodes = html_nodes(website, ".flightPageAirlineCallsign")

html_text(nodes)
```


```{, eval=F}
url2="https://www.imdb.com/search/title/?count=100&release_date=2016,2016&title_type=feature"
webpage <- read_html(url2)
rank_data_html <- html_nodes(webpage,'a')
rank_data <- html_text(rank_data_html)
rank_data
```

