---
title: "Project modelling"
author: "Jianing"
date: "15/05/2021"
output: html_document
---

---
title: "Stat442 project R scripts"
author: "Jianing Ren"
date: "05/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 0. Loading pkgs and dataset; don't run this chunk

```{r, cache=T}
library(airportr)
library(ggplot2)
library(dplyr)
library(gam)
library(leaps)
library(tidyverse)

```

```{}
#Jackson's directory 

air = read.csv("/Users/jacksonbibbens/Documents/Williams/Junior/StatLearning/Final Proj/JAN2019 entire data wo divert.csv")
plane = read.csv("/Users/jacksonbibbens/Documents/Williams/Junior/StatLearning/Final Proj/MASTER.txt")  #FAA database containing aircraft model info 
weather = read.csv("/Users/jacksonbibbens/Documents/Williams/Junior/StatLearning/Final Proj/weather.csv")
```

```{r}
#Jianing's directory

air = read.csv("/Users/rjn/OneDrive/大学/spr_2021/stat442/442project/stat442-project/data/JAN2019 entire data wo divert.csv")
plane = read.csv("/Users/rjn/OneDrive/大学/spr_2021/stat442/442project/stat442-project/data/ReleasableAircraft.2019/MASTER.txt")  #FAA database containing aircraft model info 
weather = read.csv("/Users/rjn/OneDrive/大学/spr_2021/stat442/442project/stat442-project/weather.csv")

```

# 1. Data cleaning works 
## 1a. Filtering to desired rows

```{r}
list_of_airlines = c("AA","DL","UA")
list_of_airports = c("JFK","LAX","DEN","DFW","ORD","ATL")

air_filtered = air %>% filter(OP_UNIQUE_CARRIER %in% list_of_airlines
                      & ORIGIN %in% list_of_airports
                      & DEST %in% list_of_airports)
air_filtered$TAIL_NUM = as.character(air_filtered$TAIL_NUM)
```

## 1b. Extract and merge the aircraft model info from FAA database

```{r}
plane_sub = plane %>% dplyr::select(N.NUMBER,MFR.MDL.CODE)

air_filtered$N.NUMBER = substring(air_filtered$TAIL_NUM,2,6)

air_combined = merge(air_filtered, plane_sub, by.y="N.NUMBER")  #removes flights that fails to match (either they don't have a tail number or their tail number isn't recorded in the FAA db). 
```

## 1c. Get manufacturer data

```{r}
air_combined$MFR=substring(air_combined$MFR.MDL.CODE,1,2)

#39 is Airbus; 30 is MD; 13 is Boeing
dict = data.frame(code=c(39,30,13), type=c("Airbus", "McDonnel-Douglas", "Boeing"))
air_combined$MFR.Actual = dict[match(air_combined$MFR, dict$code),]$type
```

## 1d. Route
```{r}
air_combined$route = paste0(as.character(air_combined$ORIGIN), "-",as.character(air_combined$DEST))
```


## 1f. Drop unused factor levels
```{r}
for (i in 1:ncol(air_combined))
{
    if (class(air_combined[[i]])=="factor")
        air_combined[[i]] = factor(air_combined[[i]])
}
```

## 1g. Merging Weather Data
```{r}
weather.table = data.frame(NAME = levels(weather$NAME), code=c("ATL","ORD","DFW","DEN","JFK","LAX"))
weather = merge(weather,weather.table, by.y="NAME")

origin.weather = weather[,c(19,4:8)]
names(origin.weather)[3:6] = paste0("origin_",names(origin.weather)[3:6])

dest.weather = weather[,c(19,4:8)]
names(dest.weather)[3:6] = paste0("dest_",names(dest.weather)[3:6])

air_with_weather = merge(air_combined, origin.weather, by.x=c("FL_DATE","ORIGIN"), by.y=c("DATE","code"))
air_with_weather = merge(air_with_weather, dest.weather, by.x=c("FL_DATE","DEST"), by.y=c("DATE","code"))

#write.csv(air_with_weather, "air_with_weather.csv")

rm(air_combined)
rm(air_filtered)
```


```{r}
#LAX has missing snow data; convert them to zero

ori.index = which(air_with_weather$ORIGIN=="LAX")
air_with_weather[ori.index, ]$origin_SNOW = 0

dest.index = which(air_with_weather$DEST=="LAX")
air_with_weather[dest.index,]$dest_SNOW = 0
```



## 1h. Preparing modelling data (previously 3a)
```{r}

# removing variables unsuitable for modelling 

# how to deal with the reason of delay? 

data.for.modelling = air_with_weather %>% dplyr::select(-N.NUMBER,-YEAR,-MONTH,-TAIL_NUM,-OP_CARRIER_FL_NUM,-DEP_TIME,-DEP_DELAY_GROUP,-DISTANCE_GROUP,-ARR_TIME,-CANCELLATION_CODE,-CANCELLED,-DIVERTED,-ACTUAL_ELAPSED_TIME,-FLIGHTS, -CARRIER_DELAY,-WEATHER_DELAY,-NAS_DELAY,-SECURITY_DELAY,-LATE_AIRCRAFT_DELAY,-FIRST_DEP_TIME,-TOTAL_ADD_GTIME,-LONGEST_ADD_GTIME, -X,-MFR, -WHEELS_OFF, -WHEELS_ON)


#? ONLY USE PARAMETERS MEASURED BEFORE TAKING OFF? i.e., include taxi in/out or departure delay or not
data.for.modelling = data.for.modelling %>% dplyr::select(-DEP_DELAY,-DEP_DELAY_NEW,-TAXI_IN, -TAXI_OUT)


#classifying departure and arrival time

time_breaks = c(0,900,1200,1400,1800,2100,2359)
time_labels = c("early morning","late morning","early afternoon","late afternoon","early night","late night")

data.for.modelling$CRS_DEP_TIME_GROUP = cut(data.for.modelling$CRS_DEP_TIME, breaks=time_breaks,labels=time_labels)

data.for.modelling$CRS_ARR_TIME_GROUP = cut(data.for.modelling$CRS_ARR_TIME, breaks=time_breaks,labels=time_labels)

data.for.modelling = data.for.modelling %>% dplyr::select (-CRS_ARR_TIME,-CRS_DEP_TIME)

# classifying distance


#converting day of week to actual name
data.for.modelling$DAY_OF_WEEK = as.factor(weekdays(as.Date(data.for.modelling$FL_DATE)))
data.for.modelling = data.for.modelling %>% dplyr::select (-FL_DATE) #won't be necessary given day of month

#Adding day of week binary
data.for.modelling$DAY.BIN = ifelse(data.for.modelling$DAY_OF_WEEK %in% c("Tuesday","Wednesday","Thursday"),1,0)

boxplot(ARR_DELAY~DAY.BIN, data=data.for.modelling)

```


## 1i. Further processing on missing/incomplete data
```{r}
# remove other nas

data.for.modelling = data.for.modelling[complete.cases(data.for.modelling),]


#table(test2019$MFR.MDL.CODE)
#table(train2019$MFR.MDL.CODE)
```

## ij. Creating backup before subsetting/transforming
```{r}
data.for.modelling.full = data.for.modelling #unsubsetted original; creating backup

```


## 1j. subset into delays <= 120 mins, and transformation

```{r}
#data.for.modelling = data.for.modelling.full      #restoring backup if need to undo subset
#data.for.modelling = data.for.modelling %>% filter(ARR_DELAY <= 120)     #subsetted 


min.delay = min(data.for.modelling$ARR_DELAY)

transformed.delay = log(data.for.modelling$ARR_DELAY - min.delay + 1)

original.delay = data.for.modelling.full$ARR_DELAY

hist(transformed.delay)
hist(original.delay)

qqnorm(transformed.delay);qqline(transformed.delay)
qqnorm(original.delay); qqline(original.delay)

data.for.modelling$ARR_DELAY = transformed.delay

#data.for.modelling = data.for.modelling.full #undoing subset and transformation

```




```{r worstDelays}
worst.index = which(air_with_weather$ARR_DELAY >=400 )
air_worst = air_with_weather[worst.index,]
# See lots are on Jan 19, 22 as the snowstorm came in
```

# 2 EDAs: NOW BASED SOLELY ON SUBSETTED

## 2a. Histogram and boxplots on ARR_DELAY against discrete variables

```{r}
univariate = ggplot(data.for.modelling, aes(x=ARR_DELAY))+theme_bw()


# untransformed hist


ggplot(data.for.modelling.full, aes(x=ARR_DELAY))+theme_bw()+ geom_histogram( fill="lightblue", color="black") +labs(x="Arrival Delay") +ggtitle("Untransformed Distribution of Arrival Delay")+
  theme(plot.title = element_text(hjust = 0.5, size = 20))

#ggsave("imgs/transformed_subsetted_hist.png")


# transformed histogram  
univariate + geom_histogram( fill="lightblue", color="black") +labs(x="Arrival Delay") +ggtitle("Transformed Distribution of Arrival Delay")+
  theme(plot.title = element_text(hjust = 0.5, size = 20))

#ggsave("imgs/transformed_subsetted_hist.png")

# boxplots: 

# origin
univariate + geom_boxplot(aes(y = ORIGIN), fill="lightblue")+labs(x="Arrival Delay", y="Origin")

# destination
univariate + geom_boxplot(aes(y = DEST), fill="lightblue")+labs(x="Arrival Delay", y="Destination")

# aircraft manufacturer
univariate + geom_boxplot(aes(y = MFR.Actual), fill="lightblue")+labs(x="Arrival Delay", y="Aircraft Manufacturer")+ggtitle("Transformed Arrival Delay, Distributed by Aircraft Manufacturer")+theme(plot.title = element_text(hjust = 0.5, size = 20))
#ggsave("imgs/boxplot_mfr_unsubsetted.png", width=13, height=7)

# carrier

univariate + geom_boxplot(aes(y = OP_UNIQUE_CARRIER), fill="lightblue")+labs(x="Arrival Delay", y="Airlines")+ggtitle("Transformed Arrival Delay, Distributed by Airlines")+theme(plot.title = element_text(hjust = 0.5, size = 20))
#ggsave("imgs/boxplot_airlines_unsubsetted.png", width=13, height=7)

# route (too messy; probably won't include)
ggplot(air_with_weather, aes(x=ARR_DELAY)) + geom_boxplot(aes(y=route), fill="lightblue")+labs(x="Arrival Delay", y="Route") + theme_bw()

# day of week
univariate +  geom_boxplot(aes(y = DAY_OF_WEEK), fill="lightblue") + labs(x="Arrival Delay", y="Day of Week")+ggtitle("Transformed Arrival Delay, Distributed by Day of Week")+theme(plot.title = element_text(hjust = 0.5, size = 20))
#ggsave("imgs/boxplot_dayofweek_unsubsetted.png", width=13, height=7)

# departure time group
univariate + geom_boxplot(aes(y = CRS_DEP_TIME_GROUP), fill="lightblue") + labs(x="Arrival Delay", y="Departure Time")+ggtitle("Transformed Arrival Delay, Distributed by Airlines")+theme(plot.title = element_text(hjust = 0.5, size = 20))
#ggsave("imgs/boxplot_depgroup_unsubsetted.png", width=13, height=7)


```

## 2b Scatterplot on numeric variables 
```{r}

#q.vars = data.for.modelling %>% dplyr::select(ARR_DELAY, DAY_OF_MONTH, CRS_ELAPSED_TIME, AIR_TIME, DISTANCE, origin_AWND, origin_PRCP, origin_SNOW, origin_TAVG, dest_AWND, dest_PRCP, dest_SNOW, dest_TAVG) %>% gather (-ARR_DELAY, key="var", value="value")

# main scatterplot
q.vars = data.for.modelling %>%  gather (-ARR_DELAY, -DEST, -ORIGIN, -DAY_OF_WEEK,-OP_UNIQUE_CARRIER, -MFR.Actual, -CRS_ARR_TIME_GROUP, -CRS_DEP_TIME_GROUP, -DAY.BIN, key="var", value="value")

scatter = ggplot(q.vars, aes(x=value, y=ARR_DELAY)) + facet_wrap(.~var, scales = "free") +geom_point(alpha=0.6)+labs(y="Arrival Delay") + geom_smooth(color="#06E8FF")

scatter

# scatterplot on numeric variables color coded by...

# 1. Manufacturer
ggplot(q.vars, aes(x=value, y=ARR_DELAY)) + 
    facet_wrap(.~var, scales = "free") + 
    geom_point(aes(color=MFR.Actual))+
    labs(y="Arrival Delay", color="Manufacturer", linetype="Manufacturer Fitted Lines") + 
    geom_smooth(aes(linetype=MFR.Actual), color="black", alpha=0.6)+
    scale_color_manual(values=c("Boeing"="#76e3df", "Airbus"="#aca1f0", "McDonnel-Douglas"="#fad975"))+
    scale_linetype_manual(values=c("Boeing"="solid", "Airbus"="dashed", "McDonnel-Douglas"="dotdash"))+
    ggtitle("Transformed Arrival Delay Against All Quantitative Variables")+theme(plot.title = element_text(hjust = 0.5, size = 20))+
    theme_bw()

ggsave("imgs/scatterplot_mfr_unsubsetted.png", width=13, height=7)


# 2. MD vs A+B
q.vars$MFR.BIN = ifelse(q.vars$MFR.Actual == "McDonnel-Douglas", T,F)

ggplot(q.vars, aes(x=value, y=ARR_DELAY)) + 
    facet_wrap(.~var, scales = "free") +
    geom_point(aes(color=MFR.BIN), alpha=0.6)+
    geom_smooth(aes(linetype=MFR.BIN))+
    labs(y="Arrival Delay", linetype="Is the aircraft made by McDonnel-Douglas?",color="Is the aircraft made by McDonnel-Douglas?")+
    scale_linetype()

# 3. Airlines?
ggplot(q.vars, aes(x=value, y=ARR_DELAY)) + 
    facet_wrap(.~var, scales = "free") +
    geom_point(aes(color=OP_UNIQUE_CARRIER), alpha=0.6)+
    geom_smooth(aes(linetype=OP_UNIQUE_CARRIER), color="black")+
    labs(y="Arrival Delay", linetype="Airline Fitted Lines",color="Airline")+
    scale_color_manual(values=c("AA"="#76e3df", "DL"="#aca1f0", "UA"="#fad975"))+
    scale_linetype_manual(values=c("UA"="solid", "DL"="dashed", "AA"="dotdash"))+
    ggtitle("Transformed Arrival Delay Against All Quantitative Variables")+theme(plot.title = element_text(hjust = 0.5, size = 20))+
    theme_bw()

#ggsave("imgs/scatterplot_airlines_unsubsetted.png", width=13, height=7)

```


# 3. Modelling 

## 3a. Split

```{r}
#split training and testing
set.seed(1)
randvec = sample.int(nrow(data.for.modelling), 8000)
data.for.modelling = data.for.modelling %>% dplyr::select(-route, -MFR.MDL.CODE)  #removing route and model code from data.for.modelling
train2019 = data.for.modelling[randvec,]
test2019 = data.for.modelling[-randvec,]
```


## 3b. LM
```{r}
get.test.mse = function(model.pred)
{
    mean((model.pred - test2019$ARR_DELAY)^2)
}


#AIR_TIME is probably perfectly correlated with flight distance and CRS_ELAPSED_TIME, therefore only keep one of them. If include more than one of them, R^2 will be close to 1

#Keep just the departure time group because arrival time group is correlated with departure and air time


lm1 = lm(ARR_DELAY~.-CRS_ELAPSED_TIME-AIR_TIME-CRS_ARR_TIME_GROUP-DAY.BIN, data=train2019)

summary(lm1)

pred.lm1 = predict(lm1, test2019)

mse.full.mdl = get.test.mse(pred.lm1)

#Super simplified (how did you decide which to drop? I don't think we should be dropping all high p valued predictors at once)
lm2= lm(ARR_DELAY~.-CRS_ELAPSED_TIME-AIR_TIME-MFR.Actual-DAY_OF_WEEK-ORIGIN-
          DEST-CRS_DEP_TIME_GROUP-CRS_ARR_TIME_GROUP-OP_UNIQUE_CARRIER-DAY_OF_MONTH, data=train2019)

summary(lm2)

pred.lm2 = predict(lm2, test2019)
mse.small.mdl = get.test.mse(pred.lm2)

#Jianing's backwards stepwise model 

lm.bwd.step = step(lm1, trace=0, direction="backward")

pred.bwd.step = predict(lm.bwd.step, test2019)
mse.bwd = get.test.mse(pred.bwd.step)

# forward dplyr::selection; consider lm1 as the biggest
lm.null = lm(ARR_DELAY~1, data=train2019)
lm.fwd.step = step(lm.null, scope=formula(lm1), direction = "forward", trace=0)
summary(lm.fwd.step)

pred.fwd.step = predict(lm.fwd.step, test2019)
mse.fwd = get.test.mse(pred.fwd.step)


list(mse.full.mdl, mse.small.mdl, mse.bwd, mse.fwd)
```

```{r}
# step lm assumptions and comparisons
anova(lm.bwd.step, lm1)

assumptions = function(model)
{
    par(mfrow=c(1,3))
    hist(residuals(model))
    plot(residuals(model)~model$fitted.value)
    qqnorm(residuals(model));qqline(residuals(model))
}

assumptions(lm.bwd.step)
```


## 3c LM with interaction and step

```{r}

lm.complex = update(lm1, .~.+origin_AWND*origin_SNOW+origin_AWND*origin_PRCP+origin_AWND*origin_TAVG+origin_SNOW*origin_PRCP+origin_SNOW*origin_TAVG+origin_PRCP*origin_TAVG) #six interactions between origin weathers

lm.complex = update(lm.complex, .~.+dest_AWND*dest_SNOW+dest_AWND*dest_PRCP+dest_AWND*dest_TAVG+dest_SNOW*dest_PRCP+dest_SNOW*dest_TAVG+dest_PRCP*dest_TAVG)

lm.interaction = step(lm.complex, direction="backward",trace=0)
summary(lm.interaction)

pred.interaction = predict(lm.interaction, test2019)
get.test.mse(pred.interaction)  #probably overfitting

```

## 3d. regsubsets [NOT WORKING]

```{r}
# not working; "really big"? 

#lm.step = regsubsets(ARR_DELAY~.-CRS_ELAPSED_TIME-AIR_TIME-MFR.Actual, data=train2019, nvmax=2)
```

## 3e. lasso
```{r}
library(glmnet)
set.seed(1)

lambda.grid = 10^seq(-5,5, length.out=50)
train.matrix = model.matrix(ARR_DELAY~.-CRS_ELAPSED_TIME-AIR_TIME-CRS_ARR_TIME_GROUP-DAY.BIN, data=train2019)[,-1]

test.matrix = model.matrix(ARR_DELAY~.-CRS_ELAPSED_TIME-AIR_TIME-CRS_ARR_TIME_GROUP-DAY.BIN, data=test2019)[,-1]

air.lasso = cv.glmnet(train.matrix, train2019$ARR_DELAY, alpha=1, lambda=lambda.grid, nfolds=10)

plot(air.lasso)
coef.min = coef(air.lasso, s="lambda.min")

lasso.pred = predict(air.lasso, newx = test.matrix, s="lambda.min")
get.test.mse(lasso.pred)
lasso.pred.1se = predict(air.lasso, newx = test.matrix, s="lambda.1se")
get.test.mse(lasso.pred.1se)
```

## 3f. ridge (for correlated predictors)

```{r}
set.seed(1)

air.ridge = cv.glmnet(train.matrix, train2019$ARR_DELAY, alpha=0, lambda=lambda.grid, nfolds=10)

plot(air.ridge)
coef.min = coef(air.ridge, s="lambda.min")

ridge.pred = predict(air.ridge, newx = test.matrix, s="lambda.min")
mean((ridge.pred - test2019$ARR_DELAY)^2)


coef.1se = coef(air.ridge, s="lambda.1se")

ridge.pred.1se = predict(air.ridge, newx = test.matrix, s="lambda.1se")
mean((ridge.pred.1se - test2019$ARR_DELAY)^2)

```


## 3g. elastic net

```{r}

air.net = cv.glmnet(train.matrix, train2019$ARR_DELAY, alpha=0.5, lambda=lambda.grid, nfolds=10)

plot(air.net)
coef.min = coef(air.net, s="lambda.min")

net.pred = predict(air.net, newx = test.matrix, s="lambda.min")
mean((net.pred - test2019$ARR_DELAY)^2)


coef.1se = coef(air.net, s="lambda.1se")

net.pred.1se = predict(air.net, newx = test.matrix, s="lambda.1se")
mean((net.pred.1se - test2019$ARR_DELAY)^2)

```


## 3h. GAM

```{r,warning=FALSE}
set.seed(1)

library(gam)
gam.train2019 = data.for.modelling

#gam.train2019$MFR.BIN = as.factor(gam.train2019$MFR.BIN)
gam.train2019$DAY.BIN = as.factor(gam.train2019$DAY.BIN)
gam.train2019$DEST = as.factor(gam.train2019$DEST)
gam.train2019$ORIGIN = as.factor(gam.train2019$ORIGIN)
gam.train2019$OP_UNIQUE_CARRIER = as.factor(gam.train2019$OP_UNIQUE_CARRIER)
gam.train2019$CRS_DEP_TIME_GROUP = as.factor(gam.train2019$CRS_DEP_TIME_GROUP)
gam.train2019$CRS_ARR_TIME_GROUP = as.factor(gam.train2019$CRS_ARR_TIME_GROUP)

gam.train2019 = gam.train2019 %>% dplyr::select(-CRS_ELAPSED_TIME,-AIR_TIME,-CRS_ARR_TIME_GROUP,-DAY.BIN)

gamvec = sample.int(nrow(gam.train2019), 8000)

gam.test2019 = gam.train2019[-randvec,]
gam.train2019 = gam.train2019[randvec,]


gam.start  = gam(ARR_DELAY~., data=gam.train2019)
air.scope = gam.scope(gam.train2019, response = 6)
air.scope
gam.step = step.Gam(gam.start, air.scope, direction="both",trace=0) # this isnt working...idk why 
summary(gam.step) 


anova(gam.start,gam.step)

gam.predict = predict(gam.step, gam.test2019)
mean((gam.predict-gam.test2019$ARR_DELAY)^2)

plot(gam.step)

#Assumptions for gam
assumptions(gam.step)
```


## 3i. Classification Tree

```{r}
library(tree)
tree2019 = train2019
tree2019$ARR.BIN = ifelse(tree2019$ARR_DELAY > log(76), 1,0)

tree.train = tree2019[randvec,]
tree.train = tree2019[-randvec, ]

tree.train$ARR.BIN = ifelse(tree.train$ARR_DELAY > log(76),1,0)


class.tree = tree(as.factor(ARR.BIN)~.-ARR_DELAY-CRS_ELAPSED_TIME-AIR_TIME,data= tree.train)
plot(class.tree)
text(class.tree,pretty=0)

cv.for.tree = cv.tree(class.tree, K=10, FUN=prune.tree)
index = which.min(cv.for.tree$dev)

#Not working since initial tree is so small
best.tree = prune.tree(class.tree, k=cv.for.tree$k[index]) # Prune tree
plot(best.tree)
text(best.tree,pretty=0)
```

## Regression tree
```{r}

```


# 99. Dumpster (testing or failed codes) 

### test in merging
```{, eval=F}
df1=data.frame(N.Number=c(1,3,5,6),rand=c(1000,1001,1002,1003))
df2=data.frame(N.Number=c(1,2,3,4,5), plane=c(717,727,737,747,757))
merge(df1,df2,by.y="N.Number",all.x=T)
```


### scraping

```{, eval=F}
url = paste0("https://flightaware.com/live/flight/", "N839AA")
website = read_html(url)
nodes = html_nodes(website, ".flightPageAirlineCallsign")

html_text(nodes)
```


```{, eval=F}
url2="https://www.imdb.com/search/title/?count=100&release_date=2016,2016&title_type=feature"
webpage <- read_html(url2)
rank_data_html <- html_nodes(webpage,'a')
rank_data <- html_text(rank_data_html)
rank_data
```

### Get airport coordinates (removed from 1h)

```{}

locations = as.data.frame(sapply(list_of_airports, airport_location))
locations = apply(locations, 1, unlist)
locations = data.frame(airports = rownames(locations), locations)


air_with_weather = merge(air_with_weather, locations, by.x=("ORIGIN"), by.y=("airports"))

names(air_with_weather)[c(length(names(air_with_weather))-1, length(names(air_with_weather)))] = c("origin_lat", "origin_long")

air_with_weather = merge(air_with_weather, locations, by.x=("DEST"), by.y=("airports"))

names(air_with_weather)[c(length(names(air_with_weather))-1, length(names(air_with_weather)))] = c("dest_lat", "dest_long")
```

### some exploratory plots
```{}
air_for_plotting = air_with_weather[,-c(3,4,5,6,10,11,33:41)]
air_for_plotting2 = air_for_plotting[,c(17,31:42)]

hist.subset = air_for_plotting2[air_for_plotting2$ARR_DELAY<=200,]
hist(hist.subset$ARR_DELAY, xlim=c(-100,200),breaks = seq(-100,200,by=10)) 
#Most flights have no delay

plot((ARR_DELAY)~., data=air_for_plotting2)

boxplot(ARR_DELAY~OP_UNIQUE_CARRIER, data=air_combined)

boxplot(ARR_DELAY~MFR.Actual, data=air_combined)
boxplot(ARR_DELAY~route, data=air_combined)

boxplot(ARR_DELAY~DAY_OF_WEEK,data=air_combined)
plot(ARR_DELAY~DAY_OF_MONTH,data=air_combined)

plot(ARR_DELAY~CRS_ARR_TIME, data=air_combined)

plot((ARR_DELAY)~origin_AWND, data=air_with_weather)

```

### removing unique airplane models
```{r}
#need to drop unique MFR.MDL.CODE, at least there shouldn't be new ones in testing. Therefore, filter out those plane types whose instances are fewer than 10

data.for.modelling = data.for.modelling %>% group_by(MFR.MDL.CODE) %>% filter(n()>10)
data.for.modelling$MFR.MDL.CODE = factor(data.for.modelling$MFR.MDL.CODE)
data.for.modelling = ungroup(data.for.modelling)
```

### non-ggplot EDAs

## 2a. Univariate on delay time

```{r}
#entire histogram and boxplot
boxplot(data.for.modelling$ARR_DELAY)
hist(data.for.modelling$ARR_DELAY,breaks = seq(-100,1200,by=10))

#histogram on arrival delay within 200min; this seems the best subset
hist.subset = data.for.modelling[data.for.modelling$ARR_DELAY<=200,]
hist(hist.subset$ARR_DELAY, xlim=c(-100,200),breaks = seq(-100,200,by=10)) 

#histogram on arrival delay within 400min
hist.subset = data.for.modelling[data.for.modelling$ARR_DELAY<=400,]
hist(hist.subset$ARR_DELAY, xlim=c(-100,400),breaks = seq(-100,400,by=10)) 

```

## 2b. Delay time against others
```{r, warning=F}
#categorical predictor

boxplot(ARR_DELAY~ORIGIN, data=data.for.modelling)
boxplot(ARR_DELAY~DEST, data=data.for.modelling)
boxplot(ARR_DELAY~OP_UNIQUE_CARRIER,data=data.for.modelling)
boxplot(ARR_DELAY~MFR.Actual, data=data.for.modelling)
boxplot(ARR_DELAY~route, data=air_with_weather) #route info isn't in data.for.modelling
boxplot(ARR_DELAY~DAY_OF_WEEK,data=data.for.modelling)
boxplot(ARR_DELAY~CRS_ARR_TIME_GROUP, data=data.for.modelling)
boxplot(ARR_DELAY~CRS_DEP_TIME_GROUP, data=data.for.modelling)
boxplot(ARR_DELAY~DAY.BIN, data=data.for.modelling)


#quantitative predictor

plot(ARR_DELAY~DAY_OF_MONTH, data=data.for.modelling)
plot(ARR_DELAY~CRS_ELAPSED_TIME, data=data.for.modelling)
plot(ARR_DELAY~DISTANCE, data=data.for.modelling)

#weather
plot(ARR_DELAY~., data=data.for.modelling[,c(6,13:20)]) #all weather stuff

```


```{r MFR}
airbus = air_combined[which(air_combined$MFR.Actual=="Airbus"),]
summary(air_combined[which(air_combined$MFR.Actual=="Airbus"),]$ARR_DELAY)
nrow(air_combined[which(air_combined$MFR.Actual=="Airbus"),])
hist(airbus$ARR_DELAY)

boeing = air_combined[which(air_combined$MFR.Actual=="Boeing"),]
summary(air_combined[air_combined$MFR.Actual=="Boeing",]$ARR_DELAY)
nrow(air_combined[which(air_combined$MFR.Actual=="Boeing"),])
hist(boeing$ARR_DELAY)

douglas = air_combined[which(air_combined$MFR.Actual=="McDonnel-Douglas"),]
summary(douglas$ARR_DELAY)
nrow(douglas)
hist(douglas$ARR_DELAY)

t.test(boeing$ARR_DELAY,airbus$ARR_DELAY)

t.test(boeing$ARR_DELAY,douglas$ARR_DELAY)

t.test(airbus$ARR_DELAY, douglas$ARR_DELAY)

boxplot(ARR_DELAY~MFR.MDL.CODE, data=air_combined)
summary(air_combined$MFR.MDL.CODE)

#These are the "significant" ones in the LM - all are McDonnal Douglas planes
# I think we can just look at MFR rather than Model number
special1 = air_combined[air_combined$MFR.MDL.CODE==3023688,]
special2 = air_combined[air_combined$MFR.MDL.CODE==3023689,]
special3 = air_combined[air_combined$MFR.MDL.CODE==3023690,]

#air_with_weather$MFR.BIN = as.factor(ifelse(air_with_weather$MFR.Actual=="McDonnel-Douglas",0,1))

# I think we can just keep the MFR_Actual variable, since Boeing and Airbus could also be different
```


